Conservation of interaction partners between related plants varies widely across communities and between plant families.

### 2019-05-23 manuscript/phylogenetic_signal_manuscript.tex
checked bib, ready to send out revamped discussion

\author{Alyssa R. Cirtwill$^{1,2}$, Giulio V. Dalla Riva$^{3}$, Nick J. Baker$^{1}$,\\
Mikael Ohlsson$^{4}$, Isabelle Norstr\"{o}m$^{4}$, Inger-Marie Wohlfarth Hasle$^{4}$,\\
Joshua A. Thia$^{1,5}$, Christie J. Webber$^{1}$, Daniel B. Stouffer$^{1}$}
\date{\small$^1$Centre for Integrative Ecology, School of Biological Sciences\\
\medskip$^2$Department of Ecology,\\
Environment, and Plant Sciences (DEEP)\\
Stockholm University\\
114 19 Stockholm, Sweden\\
\medskip$^3$Biomathematics Research Centre, School of Mathematics and Statistics\\
University of Canterbury\\Private Bag 4800\\
Christchurch 8140, New Zealand\\
\medskip$^4$Department of Physics, Chemistry, and Biology (IFM)\\ Link\"{o}ping University\\ 581 83 Link\"{o}ping, Sweden
\medskip$^5$Present Address: School of Biological Sciences\\
University of Queensland\\Brisbane, QLD 4072, Australia }

@@

@@

  Interactions with animals affect plants' life cycles in several critical
  ways~\citep{Mayr2001,Sauve2015}. On one hand,
  pollination and other mutualistic interactions contribute
  to the reproductive success of many angiosperms~\citep{Ollerton2011}. 
  On the other, herbivores consume plant tissues~\citep{McCall2006} which
  costs plants energy and likely lowers their fitness.
  In both cases, these interactions do not occur randomly but
  are strongly influenced by plants' phenotypes. For example, plants that 
  produce abundant or high-quality nectar may receive more visits from
  pollinators~\citep{Robertson1999} whereas plants that produce noxious 
  secondary metabolites may suffer fewer herbivores~\citep{Johnson2014}. 
  A plant's traits are also likely to determine which specific pollinators 
  and herbivores interact with that plant. Plants with different defences 
  (e.g., thorns vs. chemical defences) may deter different groups of 
  herbivores~\citep{Ehrlich1964,Johnson2014}, and the concept of 
  pollination syndromes has often been used to group plants into phenotypic
  classes believed to attract certain groups of pollinators~\citep{Waser1996,Fenster2004,Ollerton2009}.

  [[update refs, perhaps expand?]]
  If attractive and/or defensive traits are heritable,
  then we can reasonably expect that related plants will have similar 
  patterns of interactions with animals~\citep{Schemske1999}. That is, there may be \emph{phylogenetic signal} in plants' interactions such that closely-related plants may tend to have similar interaction partners.
  Recent studies that have investigated this question at the level of whole
  communities, however, have yielded mixed results~\citep{Rezende2007a,Gomez2010,Rohr2014a,Fontaine2015,Lind2015}.
  In particular, significant phylogenetic signal in plants' sets of interaction partners tends to be rare in empirical networks (\citealp{Rezende2007a,Lind2015}; but see~\citealp{Elias2013,Fontaine2015}).
  % Further, plants' roles [[how defined?]] within networks tend to be less phylogenetically constrained than those of 
  % animals~\citep{Rezende2007a,Chamberlain2014,Rohr2014,Vamosi2014,Lind2015}. [[GVDR is skeptical of this - check partiteness, size, type of role and add more detail to support.]] % I doubt this is really important, probably more trouble than it's worth.

  Several mechanisms that might weaken the conservation of interactions
  have been identified in the literature. Pollination and herbivory may be affected
  by a wide variety of traits, and not all of these are likely to be
  phylogenetically conserved~\citep{Rezende2007,Kursar2009}. If, for example,
  floral displays are strongly affected by environmental conditions~\citep{Canto2004}, 
  then pollinators may not be predicted by plants' phylogenies.
  Even if the traits affecting pollination and herbivory are
  heritable, plants may experience conflicting selection pressures that
  weaken the overall association between plant phylogeny and interaction
  partners~\citep{Armbruster1997,Lankau2007,Siepielski2010,Wise2013}. 
  For instance, floral traits that
  are attractive to pollinators can also increase 
  herbivory~\citep{Strauss2002,Adler2004,Theis2006}. 
  Conversely, herbivory can reduce pollination by inducing chemical 
  defences~\citep{Adler2006} or altering floral display or nectar 
  availability~\citep{Strauss1997}. Observed patterns
  of similarity in plants' interaction partners therefore represent
  a mixture of environmental effects and various selection pressures as
  well as plants' shared phylogenetic history.

  A further complication is the possibility that the relationship between
  plants' relatedness and the similarity of their interaction partners is
  not constant across plant clades. Closely-related plants
  in one clade might be under strong selection to favour dissimilar
  sets of pollinators to avoid exchanging pollen with other
  species~\citep{Levin1970,Bell2005,Mitchell2009}. Similar pressures 
  could also affect related plants' defences against herbivores if 
  congeners tend to grow in the same places such that herbivores 
  could easily move between them.  Unrelated plants might also 
  converge upon similar phenotypes, attracting a particularly 
  efficient or abundant pollinator~\citep{Ollerton1996,Ollerton2009}. 
  Likewise, herbivores may be able to depredate sets of unrelated 
  plants if the plants have evolved similar defences~\citep{Pichersky2000}. 
  In either case, dissimilarity of interactions among related species 
  or similarity of interactions among unrelated species could result 
  in weaker phylogenetic signal across an entire plant community. 
  Moreover, all of the aforementioned hypotheses 
  are non-exclusive; different processes likely affect different
  clades, and these processes might be associated with different 
  pressures imposed by pollination and herbivory.

  Here we use a novel approach to investigate how overlap in 
  interaction partners between pairs of plants (henceforth ``niche 
  overlap'') varies over phylogenetic distance. Whereas previous 
  studies have focused on the presence or absence of phylogenetic
  signal across entire networks, we are able to investigate the
  relationship between niche overlap and phylogenetic distance in
  within networks as well as different plant families. Specifically,
  we test whether niche overlap decreases over increasing phylogenetic
  distance in a large dataset of pollination and herbivory networks, 
  whether the plant family composition of a community affects the
  relationship between niche overlap and phylogenetic distance in that 
  community, and whether the relationship between niche overlap and 
  phylogenetic distance differs across plant families.

@@

  We studied tested for phylogenetic signal in niche overlap within a 
  set of 59 pollination and 11 herbivory networks. These networks span 
  a range of biomes (desert to scrub forest to grassland) and 
  countries (Sweden to Australia). The herbivory networks included a 
  variety of types of herbivores but were dominated by insects 
  consuming leaves. To ensure that we were analysing interactions 
  influenced by similar sets of traits across networks, we restricted 
  our herbivory networks to insects consuming leaves and excluded 
  sap-sucking, leaf-mining, and galling insects as well as seed 
  predators and xylophagous insects; all of these interactions involve 
  different plant tissues and means of feeding than leaf consumption 
  and so may be influenced by different plant and insect traits. 
  Specifically, we removed any non-leaf consuming insects and any 
  plants which had no interaction partners after removing other types  
  of herbivores. The adjusted networks range in size between 19 and 
  997 total species (mean=162, median=97) with  between 8 and 132 
  plant species (mean=39.1, median=29.5). See \emph{Table S1, 
  Supporting Information} for details on the original sources of all 
  networks.

@@

    \begin{equation}
      \omega_{ij} \propto \beta_{distance} \delta_{ij} + \beta_{pollination} I_ij + \beta_{distance:pollination} \delta_{ij} I_ij ,
    \end{equation}

@@

  To test whether the relationship between phylogenetic distance and niche overlap depended on network size, we fit a general linear model for the slope of this relationship in the glm models against the number of plant pairs for which distances could be calculated (hereafter ``network size"), network type (again using herbivory networks as a baseline), and their interaction.

@@

  The PERMANOVA tests whether there is an association between
  community composition and network-level patterns but does not
  give any information on \emph{which} plant families have the
  greatest effects. To address this, we supplemented the 
  PERMANOVA with three constrained correspondence analyses (CCAs)
  which placed plant families along an axis representing the
  change in log odds of sharing an interaction partner.
  A correspondence 
  analysis (CA) is similar to other multivariate
  analyses such as principal components analysis in that it
  reduces multivariate data to a set of orthogonal axes. A
  subset of axes that explain the majority of variation in 
  the data can then be interpreted to elucidate trends that
  were difficult to interpret in the full multivariate space.
  A constrained correspondence analysis (CCA) creates an extra
  axis based on some constraint - in this case, the change in
  log odds of sharing an interaction partner.

@@

    Across all networks, more distantly-related plants were less likely to 
    share interaction partners ($\beta_{distance}$=-6.82, 
    $p$\textless0.001). Plants in pollination networks tended to share 
    fewer interaction partners overall, and the decrease in overlap with 
    increasing phylogenetic distance was steeper 
    ($\beta_{pollination}$=-1.44, $p$\textless0.001 and 
    $\beta_{distance:pollination}$=-18.5, $p$\textless0.001, respectively). 
    That is, a pair of plants in the same genus was more likely to share 
    interaction partners than a pair of plants in the same family in both 
    types of networks, but a pair of congeners would be less likely to 
    share pollinators than to share herbivores. [[Add predicted overlap values at extremes of distance?]] These trends may be partly 
    due to the greater proportion of specialist pollinators than specialist 
    herbivores. In our dataset, an average of 48\% (+/- 14) of pollinators 
    in a given web were extreme specialists (i.e., visited only one plant 
    species) compared to 29\% (+/- 29) of herbivores ($z$=5.62, df=68, 
    $P$\textless0.001 for a binomial regression of specialists and 
    generalists over network type).

    Despite these general trends, there was substantial variation between 
    pollination networks, with overlap of interaction partners decreasing 
    with increasing phylogenetic distance in some networks and increasing 
    in others (Fig.~\ref{within_network_regression}). Overlap of interaction partners decreased significantly with increasing phylogenetic distance in 7/11 herbivory networks and 33/59 pollination networks. In the remaining four herbivory networks and 25 of the 26 remaining pollination networks, overlap of interaction partners was not related to phylogenetic distance. Overlap of interaction partners increased with increasing phylogenetic distance in only a single pollination network.% (M_PL_026).
    The slope of the relationship between phylogenetic distance and overlap of interaction partners was related to the number of plant pairs in herbivory, but not pollination, networks. Larger herbivory networks had higher values of $\beta_{distance}$ ($\beta_{size}$=2.58$\times$10$^{-4}$, $p$=0.011 for the full glm; herbivory networks are the baseline). Plant-pollinator networks had higher (less negative) slopes overall ($\beta_{pollinator}$=0.306, $p$\textless0.001 compared to the intercept value of -0.434 for herbivory networks). Plant-pollinator networks moreover showed a much weaker relationship between network size and the strength of the overlap-distance relationship ($\beta_{pollination:size}$=-2.64$\times$10$^{-4}$, $p$=0.009). After refitting the glm to the pollination networks alone, there was no significant relationship between network size and the slope of the overlap-distance relationship ($\beta_{size}$=-5.91$\times$10$^{-6}$, $p$=0.572).

    [[to SI?]]
    Comparing the results in the observed networks to those obtained after permuting phylogenetic distances across pairs of plants, the observed slope of the relationship between phylogenetic distance and interaction partner overlap was always more extreme (i.e., always lesser or always greater) than that obtained in the permuted networks (Fig.~\ref{obs_vs_random}). Observed networks with a negative relationship between phylogenetic distance and overlap always had a more negative slope than that obtained from the permuted networks, while the 10 networks with positive relationships between phylogenetic distance and overlap always had more positive relationships than the permuted networks. This suggests that even in the networks with non-significant relationships between niche overlap and phylogenetic distance, the association between niche overlap and phylogenetic distance was not random.

    [[To SI.]]
    Comparing the permuted networks to permutations of the permuted networks, the slope obtained from the initial permuted network showed no clear relationship to the slopes obtained from 500 permutations of the permuted network. Averaged over the 1000 permutations of each observed network, the slope of the permuted network was more extreme than 48.1-51.3\% of the permutations of the permuted network. This confirms that shuffling phylogenetic distances between plant pairs destroys the relationship between distance and interaction partner overlap, and that further shuffling distances does not have a predictable effect.

@@

\section*{Discussion} [[not updated post new results]]

  We found broad support for the hypothesis that more
  closely-related pairs of plants have a higher degree
  of niche overlap. Using a novel method which considers
  all pairs of plants together, 
  the probability of two plants sharing the same animal 
  interaction partners decreased with increasing 
  phylogenetic distance. Considering networks separately,
  $\approx$78\%  of the pollination and all of the 
  herbivory networks exhibited the expected trend of decreasing 
  overlap with increasing distance.

  Within families, however, there was much greater variability. 
  More than half of the plant families in each network type
  behaved as we hypothesised, with more 
  closely-related plants having greater niche overlap than 
  distantly related plants. This relationship between overlap and 
  phylogenetic distance is consistent with the idea that traits affecting 
  interactions are heritable~\citep{Schemske1999} and changing gradually
  such that closely related plants resemble their common ancestor--- and
  each other ---more than they do distantly related plants. In some families, 
  such as \emph{Asteraceae} in pollination networks, the positive slope of 
  this relationship was very shallow while in others, such as 
  \emph{Melastomataceae} in herbivory networks, the positive slope was 
  extremely steep. This could indicate different rates of phenotypic drift 
  or evolution in different families.

  In contrast, \emph{Polygonaceae} in pollination networks and 
  \emph{Fabaceae} in herbivory networks showed the opposite 
  pattern. In these families, closely-related plants had 
  \emph{lower} overlap than more distantly-related pairs of plants. 
  There are several possible reasons a plant family might
  display this pattern. First, part of the family may have recently 
  undergone a period of rapid diversification with closely-related species 
  developing novel phenotypes that attract different animal 
  interaction partners~\citep{Linder2008,Breitkopf2015}. It is also possible that the
  animals have undergone an adaptive radiation to 
  specialise on their most profitable partner~\citep{Janz2006}. Second, this pattern 
  could be the result of ecological or environmental filtering~\citep{Mayfield2009,Ackerly2010}. 
  Closely-related species which have high degrees of overlap in their interaction 
  partners might compete too severely to coexist. This is especially likely
  for plants sharing pollinators, where the loss of pollen to related species 
  might severely limit reproductive success~\citep{Levin1970,Bell2005,Mitchell2009}.
  Indeed, animal pollination and seed dispersal have been shown to act
  as filters for several plant clades~\citep{Mayfield2009}.
  Distantly-related plants with similar interaction partners,
  on the other hand, may differ in some other aspect of their niches
  and so be able to coexist. Plants sharing herbivores are unlikely to
  compete for these interaction partners, but the presence of both
  plants in a community could support higher herbivore populations
  than could one species alone~\citep{Russell2007}. If the plants 
  compete for some other resource, the combined impact of herbivory
  and competition could eliminate the rarer plant species. Distantly-related
  plants sharing herbivores, conversely, would be less likely to compete
  for vital resources and more likely to persist.

  The remaining families did not show significant relationships 
  in either direction. That is, the niche overlap between two 
  plants did not vary linearly over phylogenetic distance. Once again, there 
  are several possible drivers for this trend (or lack thereof). These plants might 
  be highly specialised on different interaction partners and therefore
  have low overlap at all levels of relatedness. In other plant families
  with more moderate levels of specialisation, it is possible 
  that pollination and/or herbivory do not exert large
  selection pressures on the plants. If traits affecting pollination
  or herbivory are not heritable in these groups~\citep{Kursar2009}
  and that their phenotypes are constrained by other factors (e.g., 
  drought tolerance), then we should not expect a relationship 
  between phylogenetic distance and overlap of interaction partners.
  Alternatively, pollination and/or herbivory might exert large 
  pressures that maintain the clade within a pollination or 
  defensive syndrome. These syndromes are commonly believed to
  predict the pollinators or herbivores with which a plant will 
  interact~\citep{Waser1996,Fenster2004,Ollerton2009,Johnson2014}.
  As some recent studies have suggested that pollination syndromes
  do not accurately predict plants' visitors in all plant 
  families~\citep{Ollerton2009},
  it may be of interest for future researchers to test whether 
  syndromes are better predictors in families with weak 
  relationships between overlap and phylogenetic distance.
  Lastly, it is possible that the absence of a linear relationship 
  between niche overlap and phylogenetic distance is because the data
  actually exhibit a strongly non-linear one. This could result, for 
  example, from an early burst of diversification followed by a period
  of stasis~\citep{Davis2014}.

>>>>

\author{Alyssa R. Cirtwill$^{1,2}$, Giulio V. Dalla Riva$^{3}$, Nick J. Baker$^{1}$,\\
Mikael Ohlsson$^{4}$, Isabelle Norstr\"{o}m$^{4}$, Inger-Marie Wohlfarth Hasle$^{4}$,\\
Joshua A. Thia$^{1,5}$, Christie J. Webber$^{1}$, Daniel B. Stouffer$^{1}$}
\date{\small$^1$Centre for Integrative Ecology, School of Biological Sciences\\
\medskip$^2$Department of Ecology,\\
Environment, and Plant Sciences (DEEP)\\
Stockholm University\\
114 19 Stockholm, Sweden\\
\medskip$^3$Biomathematics Research Centre, School of Mathematics and Statistics\\
University of Canterbury\\Private Bag 4800\\
Christchurch 8140, New Zealand\\
\medskip$^4$Department of Physics, Chemistry, and Biology (IFM)\\ Link\"{o}ping University\\ 581 83 Link\"{o}ping, Sweden\\
\medskip$^5$Present Address: School of Biological Sciences\\
University of Queensland\\Brisbane, QLD 4072, Australia }

@@

\section*{Warning}

Our original methodology was different from that in Fontaine \& Th\'{e}bault 2015 (Popul Ecol), but now that we've switched to Jaccard as our overlap measure we're doing almost the exact same thing with a slightly different set of networks. This will probably make it quite hard to argue novelty.

@@

  Interactions with animals affect plants' life cycles in several critical
  ways~\citep{Mayr2001,Sauve2016}. On one hand,
  pollination and other mutualistic interactions contribute
  to the reproductive success of many angiosperms~\citep{Ollerton2011}. 
  On the other, herbivores consume plant tissues~\citep{McCall2006} which
  costs plants energy and likely lowers their fitness.
  In both cases, these interactions do not occur randomly but
  are strongly influenced by plants' phenotypes~\citep{Fontaine2015}. 
  For example, plants that 
  produce abundant or high-quality nectar may receive more visits from
  pollinators~\citep{Robertson1999} whereas plants that produce noxious 
  secondary metabolites may suffer fewer herbivores~\citep{Johnson2014}. 
  A plant's traits are also likely to determine \emph{which} specific pollinators 
  and herbivores interact with that plant. Plants with different defences 
  (e.g., thorns vs. chemical defences) may deter different groups of 
  herbivores~\citep{Ehrlich1964,Johnson2014}, and the concept of 
  pollination syndromes has often been used to group plants into phenotypic
  classes believed to attract certain groups of pollinators~\citep{Waser1996,Fenster2004,Ollerton2009}.

  If attractive and/or defensive traits are heritable,
  then we can reasonably expect that related plants will have similar 
  patterns of interactions with animals, especially if there is some selection in either group to avoid competition or the number of potential partners is limited~\citep{Schemske1999,Ponisio2017}. That is, there may be \emph{phylogenetic signal} in plants' interactions such that closely-related plants may tend to have similar interaction partners.
  Recent studies that have investigated this question at the level of whole
  communities, however, have yielded mixed results~\citep{Rezende2007a,Gomez2010,Rohr2014a,Fontaine2015,Lind2015,Ibanez2016,Bergamini2017,Sydenham2017,Volf2017,Hutchinson2017}.
  In particular, significant phylogenetic signal in plants' sets of interaction partners tends to be rare in empirical networks (\citealp{Rezende2007a,Lind2015,Ibanez2016}; but see~\citealp{Elias2013,Fontaine2015,Hutchinson2017}). Significant degrees of phylogentic signal and/or coevolution may also lead to only small differences in network structure, adding to the difficulty of understanding patterns in species' interaction partners~\citep{Ponisio2017}.
  Further, the plant and animal components of networks can show different degrees of phylogenetic conservation of interaction partners. In mutualistic networks, animals often show a stronger phylogenetic signal in their partners than plants~\citep{Rezende2007a,Chamberlain2014,Rohr2014,Vamosi2014,Lind2015,Fontaine2015} (but see~\citep{Rafferty2013} for a counterexample). In antagonistic networks, however, actively-foraging consumers tend to show less phylogenetic signal than their prey~\citep{Ives2006,Cagnolo2011,Naisbit2011,Fontaine2015}. Thus, it is not straightforward to assume that interactions will be similar among related species. [[GVDR was skeptical of a similar point in the previous draft - less contentious now?]]
  % Further, plants' roles [[how defined?]] within networks tend to be less phylogenetically constrained than those of 
  % animals~\citep{Rezende2007a,Chamberlain2014,Rohr2014,Vamosi2014,Lind2015}. [[GVDR is skeptical of this - check partiteness, size, type of role and add more detail to support.]] % I doubt this is really important, probably more trouble than it's worth.

  Several mechanisms that might weaken the conservation of interactions
  have been identified in the literature. Pollination and herbivory may be affected
  by a wide variety of traits, and not all of these are likely to be
  phylogenetically conserved~\citep{Rezende2007,Kursar2009,Ibanez2016}. If, for example,
  floral displays are strongly affected by environmental conditions~\citep{Canto2004}, 
  then pollinators may not be predicted by plants' phylogenies.
  Even if the traits affecting pollination and herbivory are
  heritable, plants may experience conflicting selection pressures that
  weaken the overall association between plant phylogeny and interaction
  partners~\citep{Armbruster1997,Lankau2007,Siepielski2010,Wise2013,Karinho2014}. 
  For instance, floral traits that
  are attractive to pollinators can also increase 
  herbivory~\citep{Strauss2002,Adler2004,Theis2006}. 
  Conversely, herbivory can reduce pollination by inducing chemical 
  defences~\citep{Adler2006} or altering floral display or nectar 
  availability~\citep{Strauss1997}. There may also be tradeoffs between chemical and physical defenses, or defenses at different life stages~\citep{Karinho2014,Endara2017}. Observed patterns
  of similarity in plants' interaction partners therefore represent
  a mixture of environmental effects and various selection pressures as
  well as plants' shared phylogenetic history.

  A further complication is the possibility that the relationship between
  plants' relatedness and the similarity of their interaction partners is
  not constant across plant clades. Closely-related plants
  in one clade might be under strong selection to favour dissimilar
  sets of pollinators to avoid exchanging pollen with other
  species~\citep{Levin1970,Bell2005,Mitchell2009}. Similar pressures 
  could also affect related plants' defences against herbivores if 
  congeners tend to grow in the same places such that herbivores 
  could easily move between them.  Unrelated plants might also 
  converge upon similar phenotypes, attracting a particularly 
  efficient or abundant pollinator~\citep{Ollerton1996,Ollerton2009,Ibanez2016}. 
  Likewise, herbivores may be able to depredate sets of unrelated 
  plants if the plants have evolved similar defences~\citep{Pichersky2000}. 
  In either case, dissimilarity of interactions among related species 
  or similarity of interactions among unrelated species could result 
  in weaker phylogenetic signal across an entire plant community. 
  Moreover, all of the aforementioned hypotheses 
  are non-exclusive; different processes likely affect different
  clades, and these processes might be associated with different 
  pressures imposed by pollination and herbivory.

  [[Trying to emphasize the family-wise part as a point of difference from Fontaine \& Thebault, even though our results weren't the most exciting.]]
  Here we investigate how overlap in interaction partners between 
  pairs of plants (henceforth ``niche overlap'') varies over 
  phylogenetic distance and how this differs between plant families. 
  Whereas previous 
  studies have focused on the presence or absence of phylogenetic
  signal across entire networks, we take a pairwise perspective in
  order to ontain a more detailed picture of how plant phylogeny
  relates to network structure. Specifically,
  we test whether niche overlap decreases over increasing phylogenetic
  distance in a large dataset of pollination and herbivory networks, 
  whether the plant family composition of a community affects the
  relationship between niche overlap and phylogenetic distance in that 
  community, and whether the relationship between niche overlap and 
  phylogenetic distance differs across plant families.

@@

  We studied tested for phylogenetic signal in niche overlap within a 
  set of 59 pollination and 11 herbivory networks. These networks span 
  a range of biomes (desert to scrub forest to grassland) and 
  countries (Sweden to Australia). The herbivory networks included a 
  variety of types of herbivores but were dominated by insects 
  consuming leaves. To ensure that we were analysing interactions 
  influenced by similar sets of traits across networks, we restricted 
  our herbivory networks to insects consuming leaves and excluded 
  sap-sucking, leaf-mining, and galling insects as well as seed 
  predators and xylophagous insects; all of these interactions involve 
  different plant tissues and means of feeding than leaf consumption 
  and so may be influenced by different plant and insect traits. 
  Specifically, we removed any non-leaf consuming insects and any 
  plants which had no interaction partners after removing other types  
  of herbivores. The adjusted networks range in size between 19 and 
  997 total species (mean=162, median=97) with between 8 and 132 
  plant species (mean=39.1, median=29.5). See \emph{Table S1, 
  Supporting information 1} for details on the original sources of all 
  networks.

@@

    \begin{equation}
      \omega_{ij} \propto \beta_{distance} \delta_{ij} + \beta_{pollination} I_{ij} + \beta_{distance:pollination} \delta_{ij} I_{ij} ,
    \end{equation}

@@

  To test whether the relationship between phylogenetic distance and niche overlap depended on network size, we fit a general linear model for the slope of this relationship in the glm models against the number of plant pairs for which distances could be calculated (hereafter ``network size"), network type (again using herbivory networks as a baseline), and their interaction:

@@

  % # This isn't mentioned anywhere else, likely a relic of an older plan.
  % The PERMANOVA tests whether there is an association between
  % community composition and network-level patterns but does not
  % give any information on \emph{which} plant families have the
  % greatest effects. To address this, we supplemented the 
  % PERMANOVA with three constrained correspondence analyses (CCAs)
  % which placed plant families along an axis representing the
  % change in log odds of sharing an interaction partner.
  % A correspondence 
  % analysis (CA) is similar to other multivariate
  % analyses such as principal components analysis in that it
  % reduces multivariate data to a set of orthogonal axes. A
  % subset of axes that explain the majority of variation in 
  % the data can then be interpreted to elucidate trends that
  % were difficult to interpret in the full multivariate space.
  % A constrained correspondence analysis (CCA) creates an extra
  % axis based on some constraint - in this case, the change in
  % log odds of sharing an interaction partner.

@@

    Across all networks, more distantly-related plants were less likely to 
    share interaction partners ($\beta_{distance}$=-6.82, 
    $p$\textless0.001). Plants in pollination networks tended to share 
    fewer interaction partners overall, and the decrease in overlap with 
    increasing phylogenetic distance was steeper 
    ($\beta_{pollination}$=-1.44, $p$\textless0.001 and 
    $\beta_{distance:pollination}$=-18.5, $p$\textless0.001, respectively). 
    That is, a pair of plants in the same genus was more likely to share 
    interaction partners than a pair of plants in the same family in both 
    types of networks, but a pair of congeners would be less likely to 
    share pollinators than to share herbivores. As an illustration, a pair
    of plants which diverged 10mya would have a probability of 0.202 of sharing
    a given herbivore and 0.094 of sharing a given pollinator, while a 
    pair of plants which diverged 750mya would 
    have a probability of 0.121 of sharing a given herbivore or 
    0.011 of sharing a given pollinator.
    These trends may be partly 
    due to the greater proportion of specialist pollinators than specialist 
    herbivores. In our dataset, an average of 48\% (+/- 14) of pollinators 
    in a given web were extreme specialists (i.e., visited only one plant 
    species) compared to 29\% (+/- 29) of herbivores ($z$=5.62, df=68, 
    $P$\textless0.001 for a binomial regression of specialists and 
    generalists over network type). As extreme specialists by definition 
    are never shared by more than one plant, a large proportion of specialists
    would decrease interaction partner overlap.

    Despite these general trends, there was substantial variation between 
    pollination networks, with overlap of interaction partners decreasing 
    with increasing phylogenetic distance in some networks and increasing 
    in others (Fig.~\ref{within_network_regression}). Overlap of 
    interaction partners decreased significantly with increasing 
    phylogenetic distance in 7/11 herbivory networks and 33/59 
    pollination networks. In the remaining four herbivory networks and 
    25 of the 26 remaining pollination networks, overlap of interaction 
    partners was not related to phylogenetic distance. Overlap of 
    interaction partners increased with increasing phylogenetic distance 
    in only a single pollination network.% (M_PL_026).

    The slope of the relationship between phylogenetic distance and 
    overlap of interaction partners was related to the number of plant pairs 
    in herbivory, but not pollination, networks. Larger herbivory networks 
    had higher values of $\beta_{distance}$ 
    ($\beta_{size}$=2.58$\times$10$^{-4}$, $p$=0.011 for the full glm; 
    herbivory networks are the baseline). Plant-pollinator networks had 
    higher (less negative) slopes overall ($\beta_{pollinator}$=0.306, 
    $p$\textless0.001 compared to the intercept value of -0.434 for 
    herbivory networks). Plant-pollinator networks moreover showed a 
    much weaker relationship between network size and the strength of the 
    overlap-distance relationship 
    ($\beta_{pollination:size}$=-2.64$\times$10$^{-4}$, $p$=0.009). 
    After refitting the glm to the pollination networks alone, there was 
    no significant relationship between network size and the slope of 
    the overlap-distance relationship ($\beta_{size}$=-5.91$\times$10$^{-6}$, 
    $p$=0.572).

    Comparing the results in the observed networks to those obtained after 
    permuting phylogenetic distances across pairs of plants, the observed 
    slope of the relationship between phylogenetic distance and interaction 
    partner overlap was always more extreme (i.e., always lesser or always 
    greater) than that obtained in the permuted networks (Fig.~\ref{obs_vs_random}).
    Observed networks with a negative relationship between phylogenetic distance 
    and overlap always had a more negative slope than that obtained from the 
    permuted networks, while the 10 networks with positive relationships 
    between phylogenetic distance and overlap always had more positive 
    relationships than the permuted networks. This suggests that even in 
    the networks with non-significant relationships between niche overlap 
    and phylogenetic distance, the association between niche overlap and 
    phylogenetic distance was not random. When the slopes of the permuted 
    networks were compared to those obtained from permutations of the 
    permuted networks, there was no relationship (\emph{Supporting 
    information 2}).

@@

\section*{Discussion}

  We found broad support for the hypothesis that more
  closely-related pairs of plants have a higher degree
  of niche overlap. Using a novel method which considers
  all pairs of plants together, 
  the probability of two plants sharing the same animal 
  interaction partners generally decreased with increasing 
  phylogenetic distance. Considering networks separately,
  $\approx$56\%  of the pollination and $\approx$64\% of the 
  herbivory networks exhibited the expected trend of decreasing 
  overlap with increasing distance. This variation between networks
  echoes earlier studies (e.g.,~\citep{Fontaine2015,Hutchinson2017})
  which also found broad evidence for phylogenetic conservation
  of interaction partners despite variation between particular 
  networks. 
  % The lack of a relationship between phylogenetic
  % distance and interaction partner overlap in many networks 
  % could have several explanations. First, our dataset included
  % many extreme specialists, especially among the pollination 
  % networks. These species interact with only one plant and
  % therefore weaken any signal of interaction partner overlap.

  We found that the composition of plant families in a network
  was not related to the slope of the relationship between phylogenetic
  distance and interaction partner overlap. This suggests that 
  conservation of interaction partners among closely related plants
  (e.g., congeners or members of the same subfamilies) is more
  important than phylogenetic signal from deeper within the phylognetic
  tree. This echoes earlier results relating plant phylogeny to 
  predation by particular insect species~\citep{Novotny2002,Novotny2004,
  Odegaard2005} and in whole herbivory networks~\citep{Volf2017}. As
  we did not find any relationship between the families present in
  a network and the relationship between phylogenetic distance and
  interaction partner overlap in pollination or herbivory networks,
  the greater importance of shallow phylogeny may be a general 
  feature of plant-insect interaction networks.

  The variability of the relationship between partner overlap and
  phylogenetic distance at the network level and the lack of influence
  of community composition on the strength of this relationship could
  be partly due to the different trends within families. 
  More than half of the plant families in each network type
  behaved as we hypothesised, with more 
  closely-related plants having greater niche overlap than 
  distantly related plants. This relationship between overlap and 
  phylogenetic distance is consistent with the idea that traits affecting 
  interactions are heritable and change gradually
  such that closely related plants resemble their common ancestor--- and
  each other ---more than they do distantly related 
  plants~\citep{Schemske1999,Gilbert2015,Ponisio2017}. In some families, 
  such as \emph{Asteraceae} in pollination networks, the positive slope of 
  this relationship was very shallow while in others, such as 
  \emph{Melastomataceae} in herbivory networks, the positive slope was 
  extremely steep. This could indicate different rates of phenotypic drift 
  or evolution in different families. Alternatively, key traits
  affecting plant-insect interactions may not be strongly
  conserved. Studies showing a stronger relationship between trait similarity
  and shared interaction partners than phylogenetic similarity and shared
  interaction partners suggest that traits may indeed be evolutionarily
  labile~\citep{Junker2015,Ibanez2016,Endara2017}.

  Moreover, \emph{Polygonaceae} in pollination networks and 
  \emph{Fabaceae} in herbivory networks showed the opposite 
  pattern to what we expected. In these families, closely-related plants had 
  \emph{lower} overlap than more distantly-related pairs of plants. 
  There are several possible reasons a plant family might
  display this pattern. First, part of the family may have recently 
  undergone a period of rapid diversification with closely-related species 
  developing novel phenotypes that attract different animal 
  interaction partners~\citep{Linder2008,Breitkopf2015}. It is also possible that the
  animals have undergone an adaptive radiation to 
  specialise on their most profitable partner~\citep{Janz2006}. Second, this pattern 
  could be the result of ecological or environmental 
  filtering~\citep{Ackerly2003,Mayfield2009}. 
  Closely-related species which have high degrees of overlap in their interaction 
  partners might compete too severely to coexist. This is especially likely
  for plants sharing pollinators, where the loss of pollen to related species 
  might severely limit reproductive success~\citep{Levin1970,Bell2005,Mitchell2009}.
  Indeed, animal pollination and seed dispersal have been shown to act
  as filters for several plant clades~\citep{Mayfield2009}, while selection to avoid 
  competition and restriction on numbers of interaction partners may lead to
  more intimate or specialised interactions~\citep{Ponisio2017}. This is
  consistent with the relatively high proportion of extreme specialists we
  observed in the pollination networks, which likely contributes to the weak
  relationships between phylogenetic distance and interaction partner overlap
  in many networks. Finally, although we do not have information about plants'
  traits in these networks, it is possible that convergence or a high degree of
  ancestral trait conservation has occurred
  allowing distantly-related \emph{Polygonaceae} and \emph{Fabaceae} to
  interact with the same insects.
